The Name of the Title Is Hope

### 2024-12-30 paper.tex
added rabin-karp

\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsmath}
\usepackage{mathtools}

@@

\section{Background}
\subsection{The GPU Programming Environment(Model?)}
Contemporary GPUs are hierarchically organized, massively parallel processors designed to prioritize throughput over latency. Due to their specialized nature, GPUs function as co-processors to the CPU, requiring GPU programs—known as \emph{kernels} or \emph{shaders}—to be \emph{launched} or \emph{dispatched} heterogeneously by the CPU\@. Upon launch, a kernel is replicated across independent execution contexts, called threads. These threads are first organized into smaller units known as \emph{subgroups}\footnote{Also referred to as \emph{Warps}, \emph{Waves}, or \emph{SIMD-Groups}.}, which are then grouped into larger units called \emph{workgroups}\footnote{Also referred to as \emph{Cross-Thread-Arrays}, \emph{Thread-Blocks}, \emph{Blocks}, \emph{Thread-Groups}, or \emph{Groups}.}. This thread hierarchy mirrors the GPU memory hierarchy: individual threads access private, high-speed registers; threads within a workgroup share low-latency shared memory\footnote{Also referred to as \emph{Local Data Storage}, \emph{Group-Shared Memory}, (Check metal terminology).}; and all threads have access to high-bandwidth but higher-latency global memory.

@@

\subsection{The Scan Primitive}
The study of \emph{scan} (\emph{parallel prefix}) networks traces back to the design of carry-lookahead adder circuits and beyond~\cite{10.1145/322217.322232, 5219801}. A scan is typically defined on a monoid \( M \), characterized by a binary reduction operator \( \oplus \) and an identity element \( \varepsilon \). The binary operator \( \oplus \) satisfies the closure property \( \forall a, b \in M, \ (a \oplus b) \in M \) and has an identity element \( \exists \varepsilon \in M, \ \forall a \in M, \ \varepsilon \oplus a = a \). Although \( \oplus \) must be associative, it is not necessarily commutative, as demonstrated in structures like the stack monoid~\cite{}. In a scan, the result at the \( n \)-th element is the reduction of the preceding subset of elements in the sequence. If the subset includes the \( n \)-th element, it is called \textit{inclusive}; if it excludes the \( n \)-th element, it is called \textit{exclusive}. The most common scan type is the prefix sum, where \( \oplus \) is addition. For example:
\[
  x = [x_1, x_2, x_3, \dots, x_n] \ \ \ \ \ \ \ y = [1, 1, 1, 1, 1]
\]
\[
  \text{InclusiveScan}(x, \oplus) = [x_1, x_1 \oplus x_2, x_1 \oplus x_2 \oplus x_3, \dots, x_1 \oplus x_2 \oplus \cdots \oplus x_n]
\]
\[
  \text{InclusiveScan}(y, +) = [1, 2, 3, 4, 5]
\]
Due to its significance in circuits and as a fundamental algorithmic primitive, scan has been extensively studied in both electrical engineering and computer science. Harris offers a taxonomy that relates depth, fanout, and wire tracks~\cite{}, while Hinze develops an algebraic framework for scans~\cite{}. Snir proved that \emph{depth} and \emph{size} are related by (fill in)~\cite{}, and Fich proved that among minimum depth scans Ladner-Fischer networks have optimal size. Finally, Merrill and Garland provide a comprehensive review of GPU scan implementations. In this work, we focus on the evolution of inter-workgroup (global) scan strategies and evaluate their suitability for our ideal architecture.

>>>>

\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsmath}
\usepackage{mathtools}
\newcommand{\thomas}[1]{{\footnotesize\color{orange}[Thomas: #1]}}
\newcommand{\john}[1]{{\footnotesize\color{cyan}[John: #1]}}
\newcommand{\raph}[1]{{\footnotesize\color{magenta}[Raph: #1]}}

@@

\section{Background}
\subsection{The GPU Programming Environment(Model?)}
\john{So I have written, roughly, a zillion of these sections and here is my current philosophy. The danger is that you write something really interesting and relevant and a reviewer who is familiar with GPUs sees the title and sees the first few lines and says ``I can skip this'' and does. There are plenty of great descriptions of how GPUs work. Cite one of them. What this section needs to do is very specifically call out the interesting/detailed features of GPUs that are \emph{particularly relevant to this paper} that are not just ``here's generally how a GPU works''. For me, that's the last paragraph. The other part that might be necessary is defining the specific terminology used in the paper (e.g., workgroup vs.\ thread block vs.\ CTA). If you feel like breaking up a wall of text, which I generally think is a good idea, we can make a little table of equivalent terminology and boldface the terms we're using.}
Contemporary GPUs are hierarchically organized, massively parallel processors designed to prioritize throughput over latency. Due to their specialized nature, GPUs function as co-processors to the CPU, requiring GPU programs—known as \emph{kernels} or \emph{shaders}—to be \emph{launched} or \emph{dispatched} heterogeneously by the CPU\@. Upon launch, a kernel is replicated across independent execution contexts, called threads. These threads are first organized into smaller units known as \emph{subgroups}\footnote{Also referred to as \emph{Warps}, \emph{Waves}, or \emph{SIMD-Groups}.}, which are then grouped into larger units called \emph{workgroups}\footnote{Also referred to as \emph{Cross-Thread-Arrays}, \emph{Thread-Blocks}, \emph{Blocks}, \emph{Thread-Groups}, or \emph{Groups}.}. This thread hierarchy mirrors the GPU memory hierarchy: individual threads access private, high-speed registers; threads within a workgroup share low-latency shared memory\footnote{Also referred to as \emph{Local Data Storage}, \emph{Group-Shared Memory}, (Check metal terminology).}; and all threads have access to high-bandwidth but higher-latency global memory.

@@

\subsection{The Scan Primitive}
The study of \emph{scan} (\emph{parallel prefix}) networks traces back to the design of carry-lookahead adder circuits and beyond~\cite{10.1145/322217.322232, 5219801}. A scan is typically defined on a monoid \( M \), characterized by a binary reduction operator \( \oplus \) and an identity element \( \varepsilon \). The binary operator \( \oplus \) satisfies the closure property \( \forall a, b \in M, \ (a \oplus b) \in M \) and has an identity element \( \exists \varepsilon \in M, \ \forall a \in M, \ \varepsilon \oplus a = a \). Although \( \oplus \) must be associative, it is not necessarily commutative, as demonstrated in structures like the stack monoid~\cite{}. \john{Rabin-Karp string matching (``Efficient randomized pattern-matching algorithms'', Section 6, 1987)~\cite[Section 6]{Karp:1987:ERP} has 2$\times$2 matrix multiplication as its binary operation, and is a good thing to cite in a paper going to SPAA\@.} In a scan, the result at the \( n \)-th element is the reduction of the preceding subset of elements in the sequence. If the subset includes the \( n \)-th element, it is called \textit{inclusive}; if it excludes the \( n \)-th element, it is called \textit{exclusive}. The most common scan type is the prefix sum, where \( \oplus \) is addition. For example:
\[
  x = [x_1, x_2, x_3, \dots, x_n] \ \ \ \ \ \ \ y = [1, 1, 1, 1, 1]
\]
\[
  \text{InclusiveScan}(x, \oplus) = [x_1, x_1 \oplus x_2, x_1 \oplus x_2 \oplus x_3, \dots, x_1 \oplus x_2 \oplus \cdots \oplus x_n]
\]
\[
  \text{InclusiveScan}(y, +) = [1, 2, 3, 4, 5]
\]
Due to its significance in circuits and as a fundamental algorithmic primitive, scan has been extensively studied in both electrical engineering and computer science. Harris offers a taxonomy that relates depth, fanout, and wire tracks~\cite{}, while Hinze develops an algebraic framework for scans~\cite{}. Snir proved that \emph{depth} and \emph{size} are related by (fill in)~\cite{}, and Fich proved that among minimum depth scans Ladner-Fischer networks have optimal size. Finally, Merrill and Garland provide a comprehensive review of GPU scan implementations. In this work, we focus on the evolution of inter-workgroup (global) scan strategies and evaluate their suitability for our ideal architecture.
